Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-02-08 paper/structureScene.tex
Edit conclusion
answer to JAs comments

\ml{What's going on here?}
and annotation to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors. The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.

@@

From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF approaches which treat each observation the same and do not incorporate temporal structure between features. A method that takes this knowledge into account could therefore have potential for great impact in the modeling of acoustic scenes.

Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of acoustic scene similarity retrieval, much of the work dedicated to acoustic scene modeling is applied to the classification task.
An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.

The methods that performed the best all follow a similar structure, where low-level feature representations, such as MFCCs, are calculated over short windows and fed separately to a classifier, such as a support vector machine (SVM) \cite{cortes-vapnik,cristianini2000introduction}, for training. During the classification stage, a given recording is similarly processed into short windows represented by feature vectors which are classified individually by the classifier. The individial class assignments are then combined through a majority vote which assigns a class to the whole recording. This last step is known as late integration and allows the classifier to work on individual, shorter-scale sound events instead of trying to classify the recording as a whole. By doing this, the classifier is able to identify and exploit the distinct events that characterize the scene instead of trying to model its average structure. This is in agreement with the psychoacoustic results related to the perception of auditory scenes presented above.

\ml{These two last paragraphs are a bit out of place here, since we don't do classification in the paper. Perhaps there is some way to cut them down a bit?}

@@

For \emph{RbQ-w}, each scene is represented by a signature
$$
p_u = \lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace,
$$
where each of the $M$ centroids $\mu_1^u, \ldots, \mu_M^u$ are paired with corresponding weights $w_1^u, \ldots, w_M^u$. The weight $w_m^u$ for the $m$th centroid $\mu_m^u$ is the number of frames belonging to a particular cluster. The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance ($\widehat{\EMD}$) introduced by \cite{pele2008linear}. The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied with the ``ground distance'' between the histogram bins. In our case, the histogram counts are the clusters weights $w_m^u$ defined over the bins formed by the centroids $\mu_m^u$, so the ground distance is the centroid distance. \ml{From what we have below, the distance is not actually the distance between centroids, though. It's given by the kernel $K_{mn}^{uv}$.}

\ml{Do we need the details that follow? Is it not enough to refer to \cite{pele2009fast}?}
To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}. Given two signatures $p_u$ and $p_v$, the $\widehat{\EMD}$ is computed by solving the following linear program:
\begin{equation}
\begin{split}
\widehat{\EMD}(p_u,p_v) &=\left( \min\limits_{\lbrace f_{nm}\rbrace} \sum\limits_{n,m} f_{nm}D_{nm}^{uv} \right) \\
&+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace.
\end{split}
\end{equation}

\begin{equation*}
\mathrm{s.t.} \quad f_{nm}\geq0 \quad \sum\limits_{m} f_{nm} \leq w_n^u \quad \sum\limits_{n} f_{nm} \leq w_m^v
\end{equation*}

\begin{equation*}
\sum\limits_{n,m}f_{nm} = \min\left( \sum\limits_{n} w_n^u ,\sum\limits_{m} w_m^v \right)
\end{equation*}
%\ja{Can we provide some interpretation for the different terms and constraints here?}
where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin. We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$ which is computed from $K$:
\begin{equation*}
D_{mn}^{uv}=1-K_{mn}^{uv}.
\end{equation*}
\ml{Note that this is not actually a distance.}

To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernel $K^s$ is computed: %\cite{chapelle1999support,jing2003support}
\begin{equation}
\label{eq:ks}
K_{uv}^s = \exp\left( - \dfrac{\widehat{\EMD}(p_u,p_v)}{A} \right) \\
\end{equation}
with $A$ a scaling parameter, set to the mean value of the $\widehat{\EMD}$ between all the scenes. The resulting kernel $K^s$ is known as an $\EMD$ kernel, and it should be noted that there is no guarantee that this kernel is positive definite. \ml{Is this kernel used anywhere? For similarity, do we not just use the EMD distance directly?}

@@

Experiments are carried out using scattering coefficients as well as baseline mel-frequency cepstral coefficients (MFCCs) as features. For the scattering transform, each $30$-second scene is described by $128$ vectors computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$. In this case, $3$ seconds are discarded at the beginning and end of the scene to avoid boundary artifacts. Experiments are conducted with and without logarithmic compression (see Section \ref{sec:logcomp}). \ml{Results in the next section only seem to contain the logarithmic compression. Do we have the other results as well?}

@@

Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval. \ml{What does this mean?}

@@

The metric $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items. Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item. The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}. %\ja{And BOF?}
\ml{What is the early integration approach here? We should define it.}

The scaling parameter $q$ of the RBF kernels (see Eq. \ref{eq:kc}) is set to $10\%$ of the number of data points to cluster, being the number of scenes for \emph{early} and the number of clusters for RbQ approaches. \ml{What does this mean?}
The clustering for the \emph{RbQ} approaches is done using plain $k$-means with random initialization and $200$ replicates. For each method, the numbers of clusters is set to the value among $8$, $16$ and $32$ that leads to the best $p@9$ on the training set. \ml{What's the training set here?}

@@

@@

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
\label{fig:ASS_1}
\end{center}
\end{figure}

@@

This paper presents a new approach for modeling acoustic scenes which utilizes scattering transforms at small scales and a cluster-based representation at large scales. Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature. We demonstrate the usefulness of scattering-based representations for auditory scenes and the value of late integration using experiments in auditory scene classification. In order to extend the selectivity potential of late integration using SVMs to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval. For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
These outcomes shall be studied further in future work by considering larger databases and emerging tasks in computational bioacoustics \cite{wimmer2013sampling}.

>>>>

In order to identify which parts needs human inspection, one needs tools that can mine data and annotate to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors. The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.

@@

From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF approaches which treat each observation the same and do not incorporate temporal structure between features. A method that takes this knowledge into account could therefore have potential for great impact in the modeling of acoustic scenes, provided that the representation of the representation of those distinct events is rich enough.

%Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of acoustic scene similarity retrieval, much of the work dedicated to acoustic scene modeling is applied to the classification task.
%An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.

%The methods that performed the best all follow a similar structure, where low-level feature representations, such as MFCCs, are calculated over short windows and fed separately to a classifier, such as a support vector machine (SVM) \cite{cortes-vapnik,cristianini2000introduction}, for training. During the classification stage, a given recording is similarly processed into short windows represented by feature vectors which are classified individually by the classifier. The individial class assignments are then combined through a majority vote which assigns a class to the whole recording. This last step is known as late integration and allows the classifier to work on individual, shorter-scale sound events instead of trying to classify the recording as a whole. By doing this, the classifier is able to identify and exploit the distinct events that characterize the scene instead of trying to model its average structure. This is in agreement with the psychoacoustic results related to the perception of auditory scenes presented above.

%\ml{These two last paragraphs are a bit out of place here, since we don't do classification in the paper. Perhaps there is some way to cut them down a bit?}

@@

For \emph{RbQ-w}, each scene is represented by a signature
$$
p_u = \lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace,
$$
where each of the $M$ centroids $\mu_1^u, \ldots, \mu_M^u$ are paired with corresponding weights $w_1^u, \ldots, w_M^u$. The weight $w_m^u$ for the $m$th centroid $\mu_m^u$ is the number of frames belonging to a particular cluster. The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance ($\widehat{\EMD}$) introduced by \cite{pele2008linear}. The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied with the ``ground distance'' between the histogram bins. %In our case, the histogram counts are the clusters weights $w_m^u$ defined over the bins formed by the centroids $\mu_m^u$, so the ground distance is the centroid distance. \ml{From what we have below, the distance is not actually the distance between centroids, though. It's given by the kernel $K_{mn}^{uv}$.}

%\ml{Do we need the details that follow? Is it not enough to refer to \cite{pele2009fast}?}
%To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}. Given two signatures $p_u$ and $p_v$, the $\widehat{\EMD}$ is computed by solving the following linear program:
% \begin{equation}
% \begin{split}
% \widehat{\EMD}(p_u,p_v) &=\left( \min\limits_{\lbrace f_{nm}\rbrace} \sum\limits_{n,m} f_{nm}D_{nm}^{uv} \right) \\
% &+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace.
% \end{split}
% \end{equation}
%
% \begin{equation*}
% \mathrm{s.t.} \quad f_{nm}\geq0 \quad \sum\limits_{m} f_{nm} \leq w_n^u \quad \sum\limits_{n} f_{nm} \leq w_m^v
% \end{equation*}
%
% \begin{equation*}
% \sum\limits_{n,m}f_{nm} = \min\left( \sum\limits_{n} w_n^u ,\sum\limits_{m} w_m^v \right)
% \end{equation*}
% %\ja{Can we provide some interpretation for the different terms and constraints here?}
% where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin. We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$ which is computed from $K$:
% \begin{equation*}
% D_{mn}^{uv}=1-K_{mn}^{uv}.
% \end{equation*}
% \ml{Note that this is not actually a distance.}

% To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernel $K^s$ is computed: %\cite{chapelle1999support,jing2003support}
% \begin{equation}
% \label{eq:ks}
% K_{uv}^s = \exp\left( - \dfrac{\widehat{\EMD}(p_u,p_v)}{A} \right) \\
% \end{equation}
% with $A$ a scaling parameter, set to the mean value of the $\widehat{\EMD}$ between all the scenes. The resulting kernel $K^s$ is known as an $\EMD$ kernel, and it should be noted that there is no guarantee that this kernel is positive definite. \ml{Is this kernel used anywhere? For similarity, do we not just use the EMD distance directly?}

@@

Experiments are carried out using scattering coefficients as well as baseline mel-frequency cepstral coefficients (MFCCs) as features. For the scattering transform, each $30$-second scene is described by $128$ vectors computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$. In this case, $3$ seconds are discarded at the beginning and end of the scene to avoid boundary artifacts. Experiments are conducted with and without logarithmic compression (see Section \ref{sec:logcomp}).

@@

%Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval. \ml{What does this mean?}

@@

The metric $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items. Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item. The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene. The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \texit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization algorithm. Gaussian models are compared using Monte-Carlo sampling approaximation.

The scaling parameter $q$ of the RBF kernels (see Eq. \ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.%, being the number of scenes for \emph{early} and the number of clusters for RbQ approaches. \ml{What does this mean?}
The clustering for the \emph{RbQ} approaches is done using plain $k$-means with random initialization and $200$ replicates. For each method, the numbers of clusters is set to $8$.

@@

\subsubsection*{Logarithmic compression}

As can be seen on Fiugre \ref{fig:ASS_0}, the logarithmic compression is beneficial for precision performance irrespective of the similarity considered. The gain is more pronounced when considering alternative similarities.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/unsupervised_test1}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$. \ml{ob a missing ??}}
\label{fig:ASS_0}
\end{center}
\end{figure}

@@

\ja{needs to be discussed}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$. \ml{add gmms}}
\label{fig:ASS_1}
\end{center}
\end{figure}

@@

This paper presents a new approach for modeling acoustic scenes which utilizes scattering transforms at small scales and a cluster-based representation at large scales. Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature. In order to extend the selectivity potential of late integration using SVMs to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval. For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
These outcomes shall be studied further in future work by considering larger databases and emerging tasks in computational bioacoustics \cite{wimmer2013sampling}.
